An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-24 VLDB2026/main.tex
1

		\multicolumn{2}{|c|}{\textbf{Characteristics/Algorithms}} & \textbf{NHQ} & \textbf{Filtered} & \textbf{Stitched} & \textbf{CAPS} & \textbf{ACORN} & \textbf{UNG} & \textbf{Puck} & \textbf{DSG} & \textbf{iRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{Win}
		 \\ \hline
		\multirow{6}{*}{\textbf{AF}} 
		%& Filter Type         & C   & C   & C   & C   & B   & B   & B   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& AND                 & Y   & N   & N   & Y   & Y   & Y   & Y   & \multicolumn{5}{c|}{\multirow{4}{*}{\centering\shortstack{AF\\not support}}} \\ \cline{2-9}
		& OR                  & N   & Y   & Y   & N   & Y   & Y   & Y   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& Flexible Attributes & N   & Y   & Y   & Y   & Y   & Y   & Y   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& Complex Boolean     & N   & N   & N   & N   & Y   & N   & N   & \multicolumn{5}{c|}{} \\ \hline

		\multicolumn{2}{|l|}{\textbf{RF}}             & N & N & N & N & Y & N & N & Y & Y & Y & Y & Y \\ \hline
			\multicolumn{2}{|l|}{\textbf{Filter Type }}             & C & C & C & C & B & B & B & B & B & B & A/B/C & A \\ \hline
		\multicolumn{2}{|l|}{\textbf{Dynamic Insert}} & N & Y & N & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Multi Thread}}   & N & Y & Y & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Disk Support}}           &N  & Y& N & N & N & N & N & N & N & N & N & N \\ \hline
		\multicolumn{2}{|l|}{\textbf{Base}}           & Graph & Graph & Graph & IVF & Graph & Graph & IVF & Graph & Graph & Graph & Graph & Graph \\ \hline
	\end{tabular}

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				Win: WinFilter, Filtered: FilteredVamana, Stitched: StitchedVamana. 
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
				\textit{Disk Support} indicates whether the algorithm supports disk storage.
				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\end{minipage}}

@@

	\subsection{Range Filtering}

@@

	However, even though Deep and YT-Audio have similar LID values, the DSG algorithm shows large performance differences between them. This means that DSG is more sensitive to changes in data distribution. }
	%	Its robustness is lower than that of other methods, possibly because it depends not only on LID but also on other complex data characteristics.
	%	To examine the robustness of RF-ANNS algorithms, we analyze their performance across different datasets. As shown in Figure~\ref{fig:exp_8_3}, the performance differences. All methods perform worst on WIT, which has the highest LID. On Deep and YT-Audio, which have similar LID, most algorithms achieve comparable performance. 
	%	This suggests that the effectiveness of RF-ANN search is mainly influenced by LID of the dataset. A higher LID inherently increases the difficulty of hybrid query.
	%	
	%	However, the DSG algorithm shows a noticeable performance gap between the Deep and YT-Audio datasets, despite their comparable LID. This indicates that DSG is less robust than other methods when facing dataset-dependent variations, possibly due to its sensitivity to data distribution beyond LID alone.

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:Different_Computer_range}. When compared with the experimental results of the original server shown in Figure~\ref{fig:exp_8_2}a, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig:Big_Dataset_range}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:MultiModal_range} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}

	% Legend
	\includegraphics[width=0.98\columnwidth]{figures/exp/range_legend.pdf}
	\vspace{0.2cm} % Added a little space after the legend

	% --- Subfigure (a) containing two plots ---
	\begin{subfigure}[t]{0.98\columnwidth}
		\centering
		% First plot for the shared subfigure
		\includegraphics[width=0.507\linewidth]{figures/exp/range_multimodel.pdf}
		\hfill % Pushes the two plots apart
		% Second plot for the shared subfigure
		\includegraphics[width=0.48\linewidth]{figures/exp/range_multimodel_1.pdf}

		% Shared caption for the two plots above
		\captionsetup{font=small}
		\caption{MultiModal}
		\label{fig:MultiModal_range} % A single label for the combined subfigure
	\end{subfigure}

	% --- Subfigure (b) ---
	\begin{subfigure}[t]{0.507\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
		\captionsetup{font=small}
		\caption{Different Computer}
		\label{fig:Different_Computer_range}
	\end{subfigure}
	\hfill
	% --- Subfigure (c) ---
	\begin{subfigure}[t]{0.48\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
		\captionsetup{font=small}
		\caption{Big Dataset}
		\label{fig:Big_Dataset_range}
	\end{subfigure}

	\caption{\textcolor{red}{TrinityExtension——Range}}
	\label{fig:TrinityExtension——Range}
\end{figure}

	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.

>>>>

		\multicolumn{2}{|c|}{\textbf{Characteristics/Algorithms}} & \textbf{NHQ} & \textbf{Filtered} & \textbf{Stitched} & \textbf{CAPS} & \textbf{ACORN} & \textbf{UNG} & \textbf{Puck} & \textbf{DSG} & \textbf{iRange} & \textbf{SeRF} & \textbf{UNIFY} & \textbf{Win}
		 \\ \hline
		\multirow{6}{*}{\textbf{AF}} 
		& Filter Type         & C   & C   & C   & C   & B   & B   & B   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& AND                 & Y   & N   & N   & Y   & Y   & Y   & Y   & \multicolumn{5}{c|}{\multirow{4}{*}{\centering\shortstack{AF\\not support}}} \\ \cline{2-9}
		& OR                  & N   & Y   & Y   & N   & Y   & Y   & Y   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& Flexible Attributes & N   & Y   & Y   & Y   & Y   & Y   & Y   & \multicolumn{5}{c|}{} \\ \cline{2-9}
		& Complex Boolean     & N   & N   & N   & N   & Y   & N   & N   & \multicolumn{5}{c|}{} \\ \hline

		\multicolumn{2}{|l|}{\textbf{RF}}             & N & N & N & N & Y & N & N & Y & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Dynamic Insert}} & N & Y & N & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Multi Thread}}   & N & Y & Y & Y & Y & Y & Y & N & Y & Y & Y & Y \\ \hline
		\multicolumn{2}{|l|}{\textbf{Disk Support}}           &N  & Y& N & N & N & N & N & N & N & N & N & N \\ \hline
		\multicolumn{2}{|l|}{\textbf{Base}}           & Graph & Graph & Graph & IVF & Graph & Graph & IVF & Graph & Graph & Graph & Graph & Graph \\ \hline
	\end{tabular}

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				Win: WinFilter, Filtered: Filtered-DiskANN, Stitched: StitchedVamana. 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{AND} and \textit{OR} indicate whether the algorithm supports the corresponding logical operations.
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} indicates if the algorithm supports complex expressions, such as using the NOT operator or nested logic with parentheses.
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
				\textit{Disk Support} indicates whether the algorithm supports disk storage.
				\textit{Base} indicates the foundational ANN algorithm for the hybrid query method.
		\end{minipage}}

@@

		\hspace*{5pt}
		% 放大的图例（占满单栏）
		\includegraphics[width=0.98\columnwidth]{figures/indexData/rangeFilter_legend_only.pdf}

		%\vspace{0.1cm}

		% 三图横排
		\begin{subfigure}[t]{0.495\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/indexData/rangeFilter_build_time_comparison_query.pdf}
			\caption{\footnotesize Construction time}
			\label{fig:rangeFilter_build_time}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.495\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/indexData/rangeFilter_index_size_mb_comparison_query.pdf}
			\caption{\footnotesize Index size}
			\label{fig:rangeFilter_index_size_mb}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.495\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/indexData/rangeFilter_memory_mb_comparison_query.pdf}
			\caption{\footnotesize Build peak memory}
			\label{fig:rangeFilter_build_memory_mb}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.495\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/searchMem/range_memory_comparison.pdf}
			\caption{\footnotesize Search peak memory}
			\label{fig:rangeFilter_search_memory_mb}
		\end{subfigure}

		%		\caption{Time and space overhead of range filtering index construction}
%		\caption{Time and space overhead of index construction}
		\caption{\textcolor{violet}{Time and space overhead}}
		\label{fig:rangeFilter_build_index_comparison}
	\end{figure}
	\subsection{Range Filtering}

@@

	However, even though Deep and YT-Audio have similar LID values, the DSG algorithm shows large performance differences between them. This means that DSG is more sensitive to changes in data distribution. }
	%	Its robustness is lower than that of other methods, possibly because it depends not only on LID but also on other complex data characteristics.
	%	To examine the robustness of RF-ANNS algorithms, we analyze their performance across different datasets. As shown in Figure~\ref{fig:exp_8_3}, the performance differences. All methods perform worst on WIT, which has the highest LID. On Deep and YT-Audio, which have similar LID, most algorithms achieve comparable performance. 
	%	This suggests that the effectiveness of RF-ANN search is mainly influenced by LID of the dataset. A higher LID inherently increases the difficulty of hybrid query.
	%	
	%	However, the DSG algorithm shows a noticeable performance gap between the Deep and YT-Audio datasets, despite their comparable LID. This indicates that DSG is less robust than other methods when facing dataset-dependent variations, possibly due to its sensitivity to data distribution beyond LID alone.
\begin{figure*}[t]
	\centering

@@

	\textit{\textbf{\textcolor{blue}{Different Hardware Environments.}}}
	\textcolor{cyan}{Consistent with Section 4.2.4, we conduct additional range filtering experiments on the YT-Audio-Real dataset using the same server configuration, with results presented in Figure~\ref{fig:range-cross-platform}. When compared with the experimental results of the original server shown in Figure~\ref{fig:exp_8_2}a, our findings remain consistent with those reported in Section 4.2.4. For brevity and due to space limitations, we omit further discussion of these redundant results.}
%	\textcolor{cyan}{Consistent with Section 4.2.4, we conducted additional range filtering experiments using the YT-Audio-Real dataset on the same server, with results shown in Figure \ref{fig:Different Computer range}. Compared with the original server's experimental results (Figure \ref{fig:exp_8_2}($e$)), the conclusions remain identical to those described in Section 4.2.4. Due to space constraints, these details are not reiterated here.}

	\textit{\textbf{\textcolor{red}{Large-scale Datasets.}}} \textcolor{red}{As shown in the figure~\ref{fig: range big dataset}, On the 100M dataset, the top 3 ranked algorithms (WinFilter, UNIFY, and iRange) remain unchanged. However, their high performance is achieved at the cost of extremely high memory consumption, as detailed in Figure~\ref{fig:rangeFilter_search_memory_mb}. Meanwhile, VBASE and Faiss demonstrate moderate performance but significantly lower memory consumption than the top three algorithms}.

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:range-multimodal} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

%	\textit{\textbf{Multimodel, 85 and Deep100M.}} \textcolor{red}{The fig is Figure \ref{fig:TrinityExtension——Range}}
	\section{DISCUSSION}
	Based on the performance of the algorithms on different experimental scenarios, we discuss our findings as follows.
